\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)

### 2015-09-21 methodology.tex
updated section 4

\textbf{Intercepting network messages}.
The real ExtMgr uses a network engine to asynchronously send messages to the ENs. The \psharp test harness mocks the original network engine in vNext by overriding its interface. This enables the mocked network engine (we use the terms \textit{mocked} and \textit{modeled} interchangeably) to intercept all outbound messages and relay them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines. As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}. This mocked network engine replaces the real network engine in the wrapped ExtMgr, as shown in Figure~\ref{fig:wrap_target}.

@@

// mocked engine for intercepting Extent Manager messages
class MockedNetEngine : NetworkEngine {
  public override void SendMessage(Socket s, Message msg) {
    // intercept and relay Extent Manager messages
    PSharpRuntime.Send(TestingDriver,
      new ExtentManagerMsgEvent(), s, msg);
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Mocked network engine in vNext.}
\label{fig:enginecode}
\vspace{-2mm}
\end{figure}

@@

  // extent node sync logic
  [OnEvent(TimerTickEvent, ProcessExtentNodeSync)]
  void ProcessExtentNodeSync() {
    var sync = ExtCtr.GetSyncReport(); // prepare sync report
    this.Send(ExtentManagerMachine, new ENMsgEvent(), sync);
  }

  // extent node failure logic
  [OnEvent(FailEvent, ProcessFailure)]
  void ProcessFailure() {
    // notifies the Extent Manager that this EN failed
    this.Send(ExtentManagerMachine, new ENMsgEvent(),
      new ExtentNodeDeletion(ExtentId));
    // notifies the monitor that this EN failed
    this.Monitor<RepairMonitor>(new ENFailedEvent(), this);
    this.Halt(); // terminate this P# machine
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{The modeled EN in vNext.}
\label{fig:mocked_en}
\vspace{-2mm}
\end{figure}

@@

In the modeled EN sync logic, the machine is driven by an external timer modeled in \psharp (see \S\ref{sec:method:timer}). It prepares a sync report with \texttt{extCtr.GetSyncReport(...)}, and then asynchronously sends the report to \texttt{ExtentManager} using \texttt{PSharpRuntime.Send(...)}. The \texttt{ExtentNode} machine also includes failure-related logic, which is discussed in detail in \S\ref{sec:method:driver}.

@@

\begin{figure}[t]
\begin{lstlisting}
// modeling timer expiration in P#
class Timer : Machine {
  MachineId Target; // target machine

  [OnEvent(RepeatEvent, GenerateTimeTick)]
  void GenerateTimeTick() {
    // nondeterministic choice controlled by P#
    if (this.Nondet())
      this.Send(this.Target, new TimerTickEvent());
    this.Send(this, new RepeatEvent()); // loop
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Modeling timer expiration using \psharp.}
\label{fig:mocked_timer}
\vspace{-2mm}
\end{figure}

System correctness should \emph{not} hinge on the frequency of any individual timer. Instead, all nondeterminism due to timing-related events must be delegated to \psharp. To achieve this, all timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach also used in previous work~\cite{desai2015building}. Similarly, the timers in all \texttt{ExtentNode} machines are disabled and replaced by \psharp models.

Figure~\ref{fig:mocked_timer} shows the \texttt{Timer} machine of vNext. \texttt{Timer} invokes the \psharp method \texttt{Nondet()} which generates a nondeterministic choice that is controlled by the \psharp runtime. Using \texttt{Nondet()} allows the machine to nondeterministically send a timeout event to a target machine. The \psharp testing engine has the freedom to schedule arbitrary interleavings between these timeout events and all other regular system events.

@@

  // other test harness driving logic
  ...
  void InjectNodeFailure() {
    // nondeterministically choose an EN using P#
    var node = (MachineId)this.Nondet(ExtentNodes);    
    this.Send(node, new FailEvent()); // fail the chosen EN
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{The TestingDriver machine in vNext.}
\label{fig:testing_driver}
\vspace{-2mm}
\end{figure}

The \texttt{TestingDriver} machine drives two testing scenarios. In the first scenario, \texttt{TestingDriver} launches one \texttt{ExtentManager} and three \texttt{ExtentNode} machines, with a single extent on one of the ENs. \texttt{TestingDriver} then waits for the extent to be replicated at the remaining ENs. In the second testing scenario, \texttt{TestingDriver} fails one of the \texttt{ExtentNode} machines and launches a new one. \texttt{TestingDriver} then waits for the extent to be repaired on the newly launched \texttt{ExtentNode} machine.

Figure~\ref{fig:testing_driver} illustrates how \psharp can be used for injecting nondeterministic failures. The \texttt{TestingDriver} machine uses \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} from the set of all available \texttt{ExtentNode} machines. If \texttt{n} is the chosen machine, \texttt{TestingDriver} emulates a failure by sending a \texttt{FailEvent} event to \texttt{n}. Machine \texttt{n} handles this event by wrapping the \texttt{ExtentNodeDeletion} message, which is a failure notification, to the \texttt{ENMsgEvent} and sending it to the \texttt{ExtentManager} machine (see Figure~\ref{fig:mocked_en}). \texttt{ExtentNodeDeletion} is the same message that is used by the real vNext for simulating failures in its unit tests; we simply re-use this functionality. Next, \texttt{n} notifies the liveness monitor that it has failed. Finally, machine \texttt{n} calls the \psharp method \texttt{Halt()}, which causes \texttt{n} to terminate. \psharp not only enumerates interleavings of asynchronous event handlers, but also the values returned by calls to \texttt{Nondet()}, thus enumerating different failure scenarios.

@@

\begin{figure}[t]
\begin{lstlisting}
class RepairMonitor : Monitor {
  // true: EN has replica, false: EN has no replica
  private Dictionary<Machine, bool> ExtentNode2Replica;

  // cold state: repaired
  [OnEvent(ENFailedEvent, ProcessExtentNodeFailure)]
  cold state Repaired {
    void ProcessExtentNodeFailure() {
        var node = GetExtentNode(this.Payload);
        ExtentNode2Replica.Remove(node);
        goto Repairing;
    }
  }

  // hot state: repairing
  [OnEvent(ExtentRepairedEvent, ProcessExtentRepaired)]
  hot state Repairing {
    void ProcessExtentRepaired() {
      var node = GetExtentNode(this.Payload);
      ExtentNode2Replica[node] = true;
      if (ReplicaCount == Harness.REPLICA_COUNT_TARGET)
        goto Repaired;
    }
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{The RepairMonitor liveness monitor.}
\label{fig:monitor}
\vspace{-2mm}
\end{figure}

>>>>

\textbf{Intercepting network messages}.
The real ExtMgr uses a network engine to asynchronously send messages to the ENs. The \psharp test harness mocks the original network engine in vNext by overriding its interface. This enables the mocked network engine (we use the terms \textit{mocked} and \textit{modeled} interchangeably) to intercept all outbound messages and relay them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines. As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharp.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}. This mocked network engine replaces the real network engine in the wrapped ExtMgr, as shown in Figure~\ref{fig:wrap_target}.

@@

// mocked engine for intercepting Extent Manager messages
class MockedNetEngine : NetworkEngine {
  public override void SendMessage(Socket s, Message msg) {
    // intercept and relay Extent Manager messages
    PSharp.Send(TestingDriver, new ExtMgrMsgEvent(), s, msg);
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Mocked network engine in vNext.}
\label{fig:enginecode}
\vspace{-2mm}
\end{figure}

@@

  // extent node sync logic
  [OnEvent(TimerTickEvent, ProcessExtentNodeSync)]
  void ProcessExtentNodeSync() {
    var sync = ExtCtr.GetSyncReport(); // prepare sync report
    PSharp.Send(ExtentManagerMachine, 
      new ENMsgEvent(), sync);
  }

  // extent node failure logic
  [OnEvent(FailureEvent, ProcessFailure)]
  void ProcessFailure() {
    // notifies the monitor that this EN failed
    PSharp.Monitor<RepairMonitor>(new ENFailedEvent(), this);
    PSharp.Halt(); // terminate this P# machine
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{The modeled EN in vNext.}
\label{fig:mocked_en}
\vspace{-2mm}
\end{figure}

@@

In the modeled EN sync logic, the machine is driven by an external timer modeled in \psharp (see \S\ref{sec:method:timer}). It prepares a sync report with \texttt{extCtr.GetSyncReport(...)}, and then asynchronously sends the report to \texttt{ExtentManager} using \texttt{PSharp.Send(...)}. The \texttt{ExtentNode} machine also includes failure-related logic, which is discussed in detail in \S\ref{sec:method:driver}.

@@

\begin{figure}[t]
\begin{lstlisting}
// modeling timer expiration in P#
class Timer : Machine {
  Machine Target; // target machine

  [OnEvent(RepeatedEvent, GenerateTimerTick)]
  void GenerateTimerTick() {
    // nondeterministic choice controlled by P#
    if (PSharp.Nondet())
      PSharp.Send(Target, new TimerTickEvent());
    PSharp.Send(this, new RepeatedEvent()); // loop
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{Modeling timer expiration using \psharp.}
\label{fig:mocked_timer}
\vspace{-2mm}
\end{figure}

System correctness should \emph{not} hinge on the frequency of any individual timer. Hence, it makes sense to delegate all nondeterminism due to timing-related events to \psharp. To achieve this, all the timers inside ExtMgr are disabled (see Figure~\ref{fig:wrap_target}), and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp.
%, an approach also used in previous work~\cite{desai2015building}.
Similarly, \texttt{ExtentNode} machines do {\em not} have internal timers either. Their periodic heartbeats and sync reports are also driven by timers modeled in \psharp.

Figure~\ref{fig:mocked_timer} shows the \texttt{Timer} machine in the test harness. \texttt{Timer} invokes the \psharp method \texttt{Nondet()} which generates a nondeterministic choice controlled by the \psharp runtime. Using \texttt{Nondet()} allows the machine to nondeterministically send a timeout event to its target (the \texttt{ExtentManager} or \texttt{ExtentNode} machines). The \psharp testing engine has the freedom to schedule arbitrary interleavings between these timeout events and all other regular system events.

@@

  // other test harness logic
  ...
  void InjectNodeFailure() {
    // nondeterministically choose an EN using P#
    var node = (Machine)PSharp.Nondet(ExtentNodes);    
    PSharp.Send(node, new FailureEvent()); // fail chosen EN
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{The TestingDriver machine in vNext.}
\label{fig:testing_driver}
\vspace{-2mm}
\end{figure}

The \texttt{TestingDriver} machine drives two testing scenarios. In the first scenario, \texttt{TestingDriver} launches one \texttt{ExtentManager} and three \texttt{ExtentNode} machines, with a single extent on one of the ENs. It then waits for the extent to be replicated at the remaining ENs. In the second testing scenario, \texttt{TestingDriver} fails one of the \texttt{ExtentNode} machines and launches a new one. It then waits for the extent to be repaired on the newly launched \texttt{ExtentNode} machine.

Figure~\ref{fig:testing_driver} illustrates how \texttt{TestingDriver} leverages \psharp to inject nondeterministic failures. It uses \texttt{Nondet()} to nondeterministically choose an \texttt{ExtentNode} machine and sends a \texttt{FailureEvent} to the chosen machine to emulate EN failure. As shown in the earlier Figure~\ref{fig:mocked_en}, the chosen \texttt{ExtentNode} machine processes the \texttt{FailureEvent},
%and instructs \texttt{ExtentManager} to expire itself by sending a \texttt{ExtentNodeDeletion} message.\texttt{ExtentNodeDeletion} is the same message that is used by the real \texttt{ExtMgr} to expire an EN. At the same time, \texttt{ExtentNode}
notifies the liveness monitor of its failure and terminates itself by invoking \texttt{PSharp.Halt()}. \psharp not only enumerates interleavings of asynchronous event handlers, but also the values returned by calls to \texttt{Nondet()}, thus enumerating different failure scenarios.

@@

\begin{figure}[t]
\begin{lstlisting}
class RepairMonitor : Monitor {
  private HashSet<Machine> ExtentNodesWithReplica;

  // cold state: repaired
  [OnEvent(ENFailedEvent, ProcessExtentNodeFailure)]
  cold state Repaired {
    void ProcessExtentNodeFailure() {
        var node = GetExtentNode(this.Payload);
        ExtentNodesWithReplica.Remove(node);
        if (ReplicaCount < Harness.REPLICA_COUNT_TARGET)				
          goto Repairing;
    }
  }

  // hot state: repairing
  [OnEvent(ExtentRepairedEvent, ProcessExtentRepaired)]
  hot state Repairing {
    void ProcessExtentRepaired() {
      var node = GetExtentNode(this.Payload);
      ExtentNodesWithReplica.Add(node);
      if (ReplicaCount == Harness.REPLICA_COUNT_TARGET)
        goto Repaired;
    }
  }
}
\end{lstlisting}
\vspace{-4mm}
\caption{The RepairMonitor liveness monitor.}
\label{fig:monitor}
\vspace{-2mm}
\end{figure}
